---
title: "p8105_hw3_fy2297"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
library(hexbin)
# some figure preferences in global options (copied from lecture)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
# Problem 1

```{r}
library(p8105.datasets)
data("instacart")
insta_df =
  instacart %>% 
  janitor::clean_names()
```

```{r}
# find how many aisles and which aisles are the most popular.
insta_df %>% 
  count(aisle) %>% 
  arrange(-n)
```
we can see that there are 134 aisles, with "fresh vegetables", "fresh fruits", and "packaged vegetables fruits" being the top 3 aisles people order from.

```{r}
# make a plot that shows the number of items ordered in each aisle
insta_df %>% 
  count(aisle) %>% 
  filter(n>10000) %>% 
  arrange(-n) %>% 
  mutate(aisle = fct_reorder(aisle, n, .desc = TRUE)) %>% 
  ggplot(aes(x=aisle, y=n)) + geom_point()+
  labs(
    title = "Items ordered in each aisle",
    x = "Aisle name",
    y = "Items ordered"
  ) +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r}
# make a table showing the three most popular items in required aisles
insta_df %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  arrange(-n) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank<=3) %>% 
  knitr::kable()
```

```{r}
# make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week
insta_df %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  knitr::kable(digits = 1)
```

# Problem 2
```{r}
# load, tidy, and otherwise wrangle the data.
accel_df = read_csv("./datahw3/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    values_to = "activity_value"
  ) %>% 
  mutate(
    weekday_vs_weekend = case_when(
      day == "Saturday" | day == "Sunday" ~ "weekend",
      day != "Saturday" | day != "Sunday" ~ "weekday",
      TRUE                                ~ ""
    )
  )
```

```{r echo = FALSE}
# data needed for description
num_row = nrow(accel_df)
num_col = ncol(accel_df)
```
## describe the resulting dataset
The size of the result cleaned and tidied dataset is `r num_row`*`r num_col`, and it has `r num_row` observations. There are variables like "week", "day id", "day", "activity", "activity value", and "weekday vs weekend" indicating whether the day is a weekday or weekend.

```{r}
# create a table that shows total activity for each day
 # table 1: separate by week
accel_df %>% 
  group_by(week, day) %>% 
  summarize(
    total_activity = sum(activity_value)
  ) %>%
  arrange(day) %>% 
  group_by(day) %>% 
  pivot_wider(
    names_from = day,
    values_from = total_activity
  ) %>% 
  knitr::kable(digit = 1)

 # table 2: add same day in different weeks
accel_df %>% 
  group_by(day) %>% 
  summarize(
    total_activity = sum(activity_value)
  ) %>%
  arrange(day) %>% 
  group_by(day) %>% 
  pivot_wider(
    names_from = day,
    values_from = total_activity
  ) %>% 
  knitr::kable(digit = 1)
```
## describe the trend in the table
Saturday has the lowest total activity, especially week 4 and week 5. Other than that, no obvious trend or pattern.

```{r}
# create a dataset for plotting
accel_plot_df = 
  accel_df %>% 
  group_by(day,activity) %>% 
  mutate(activity = parse_number(activity)) %>% 
  summarize(
    mean_activity_by_day = mean(activity_value)
  )

# make a plot that shows 24-hour activity time for each day
ggplot(accel_plot_df, aes(x = activity, y = mean_activity_by_day, color = day)) + 
   labs(
    title = "24-hour activity time in a day",
      x = "time in a day (24-hour)",
      y = "activity value"
  ) +
  theme(plot.title = element_text(hjust = 0.5)) +
    scale_x_continuous(
    breaks = c(0, 180, 360, 540, 720, 900, 1080, 1260, 1440),
    labels = c("0:00", "3:00", "6:00", "9:00", "12:00", "15:00", "18:00", "21:00", "23:59"),
    limits = c(0, 1440)) +
  geom_line()
```

## describe the trend in the plot
Generally, around 7:30am, 12 at noon, and 9pm, people tend to be more active. And, people are more active at noon on weekends.


# Problem 3
```{r}
# load and clean the data for future use
library(p8105.datasets)
data("ny_noaa")

noaa_df = 
  ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month", "day"), sep = '-') %>% 
  rename(
    tmax_celcius = tmax,
    tmin_celcius = tmin,
    prcp_mm = prcp,
    snow_mm = snow
  ) %>% 
  mutate(
    tmax_celcius = as.numeric(tmax_celcius)/10,
    tmin_celcius = as.numeric(tmin_celcius)/10,
    prcp_mm = prcp_mm/10,
    year = as.numeric(year)
  )
# find the most commonly observed value for snowfall
noaa_snow_count = 
  noaa_df %>% 
  group_by(snow_mm) %>% 
  summarize(
    n_obs_snow = n()
  ) %>% 
  arrange(-n_obs_snow)
```
The most commonly observed values for snowfall are 0mm and NA because NY doesn't snow for the most of time, and maybe sometimes when it doesn't snow, the data isn't recorded, the data makes sense.

```{r}
# make a two-panel plot showing the average max temp. in Jan and July in each station across years
temp_ave_max_plot = 
  noaa_df %>% 
  filter(month == "07" | month == "01")%>% 
  group_by(year,month,id) %>% 
  summarize(average_tmax = mean(tmax_celcius, na.rm = TRUE)) %>% 
  ggplot(aes(x =year, y = average_tmax, color = month)) + geom_point() + geom_smooth() + labs(
    title = "max temperature in January and July in each station across years",
    x = "year",
    y = "avarage max temperature (celcius)"
  ) +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_grid(. ~month)
temp_ave_max_plot
```

```{r}
# make a two-panel plot showing 1. tmax vs tmin for the full dataset 2. distribution of snowfall values greater than 0 and less than 100 separately by year.
tmax_vs_tmin = 
  noaa_df %>% 
  drop_na(tmax_celcius, tmin_celcius) %>% 
  ggplot(aes(x = tmin_celcius, y = tmax_celcius)) + geom_hex()+
  labs(
    title = "tmax vs tmin (celcius)",
      x = "tmin (celcius)",
      y = "tmax (celcius)"
  ) +
  theme(legend.position = "right") +
  theme(plot.title = element_text(hjust = 0.5))
tmax_vs_tmin

dist_snow = 
  noaa_df %>% 
  filter(snow_mm < 100 & snow_mm >0) %>% 
  group_by(year) %>% 
  ggplot(aes(x = year, y = snow_mm, group = year)) +
  geom_boxplot()+
  labs(
    title = "snow distribution by year (between 0mm and 100mm)",
      x = "year",
      y = "snowfall (mm)") +
   scale_x_continuous(
    breaks = c(1980, 1985, 1990, 1995, 2000, 2005, 2010)
  )+
  theme(plot.title = element_text(hjust = 0.5))
dist_snow



tmax_vs_tmin/dist_snow
```







